|

### 2023-01-09 paper.Rmd
autogit

To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates. I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each the DAE-LSTM. The reconstruction error is measured for two separate populations, a training population $train\_pop$ that is used to train our DAE-LSTM as well as a hold-out validation population $test\_pop$. For the pre-trained DAE-GP two additional, separate populations $\hat{train\_pop}$ and  $\hat{test\_pop}$ are used only for pre-training.

@@

For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=$120$). To avoid creating biased results through the presence of sampling error inside the pre-training population (see @sampling_err_gp), the population size for the pre-training phase is chosen very high with $\hat{pop\_size}=20000$ where 50% of $\hat{pop}$ is used for the training population $\hat{train\_pop}$ and the remaining 50% are used for the hold-out validation population $\hat{test\_pop}$. The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent. The reason for using a high amount of 1000 fixed training epochs is to deliberately force the DAE-LSTM to overfit to the training data.

In general I expect that with a growing number of model parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population $train\_pop$ resulting in a small reconstruction error for the training population and a large one for the validation population $train\_pop$. The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases [@weigend1994overfitting].

@@

![First Generation Median Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_median_training_error_by_neurons.png)

![First Generation final Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_neurons.png)

@@

![First Generation final Training Error for variable number of hidden Layers - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_boxplot_training_error_by_layer.png)

>>>>

To study generalization behavior this series of experiments is conducted using DAE-GP with only a single generation until the search terminates. I disregard the fitness of solutions found and focus soley on the reconstruction error that is produced during the training of each the DAE-LSTM. The reconstruction error is measured for two separate populations, a training population $P_{train}$ that is used to train our DAE-LSTM as well as a hold-out validation population $P_{test}$. For the pre-trained DAE-GP two additional, separate populations $\hat{P}_{train}$ and  $\hat{P}_{test}$ are used exclusively for pre-training.

@@

For each configuration I tested traditional DAE-GP as well as a pre-trained DAE-GP resulting in 12 total sub experiments that were each based on 10 individual runs (total number of runs=$120$). To avoid creating biased results through the presence of sampling error inside the pre-training population (see @sampling_err_gp), the population size for the pre-training phase is chosen very high with the size of $\hat{P}$ =20000 where 50% of $\hat{P}$ is used for the training population $\hat{P}_{train}$ and the remaining 50% are used for the hold-out validation population $\hat{P}_{test}$. The training of DAE-LSTM uses a fixed number of 1000 epochs until termination and uses Adam optimization for gradient descent. The reason for using a high amount of 1000 fixed training epochs is to deliberately force the DAE-LSTM to overfit to the training data.

In general I expect that with a growing number of model parameters (either by adding more hidden layers or more hidden neurons per layer) the DAE-LSTM will be more prone to overfit to the training population $P_{train}$ resulting in a small reconstruction error for the training population and a large one for the validation population $P_{test}$. The reason for this effect is that a larger network trained over an extended period of time (without the use of strategies like early stopping), has much more potential to learn noise from the training dataset than a smaller network, which is more likely to result in worsening performance for previously unseen cases [@weigend1994overfitting].

@@

![First Generation Median Training Error for variable number of hidden Neurons - Airfoil](./img/airfoil_firstGen/airfoil_firstGen_mean_training_error_by_neurons.png)

@@
